---
title: "hw7"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

Load necessary R packages and the Associated Press dataset
```{r}
library(topicmodels)
library(tm)
library(SnowballC)
library(tidytext)
library(ggplot2)
library(dplyr)
library(tidyr)
library(tibble)

data("AssociatedPress")
AssociatedPress
```

Perform stemming on the terms in the AssociatedPress dataset
```{r}
terms = Terms(AssociatedPress)
stemmed_terms = wordStem(terms, language = "english")

dimnames(AssociatedPress)$Terms = stemmed_terms
```

Apply Latent Dirichlet Allocation (LDA) model to the AssociatedPress dataset
```{r}
ap_lda <- LDA(AssociatedPress, k = 12, control = list(seed = 1234))
```

Extract posterior probabilities from the LDA model
```{r}
ap_lda <- posterior(ap_lda, newdata = AssociatedPress)$terms %>%
  as.data.frame(stringsAsFactors = FALSE) %>% 
  t() %>% 
  as.data.frame(stringsAsFactors = FALSE)

colnames(ap_lda) <- paste("topic", 1:ncol(ap_lda))

ap_topics <- ap_lda %>%
  rownames_to_column("term") %>%
  pivot_longer(cols = starts_with("topic"), 
               names_to = "topic", 
               values_to = "distinctivity")
ap_topics
```

Plot the top 10 distinct terms for each topic
```{r}
ap_top_terms <- ap_topics %>%
  group_by(topic) %>%
  slice_max(distinctivity, n = 10) %>%
  ungroup() %>%
  arrange(topic, -distinctivity)

ap_top_terms %>%
  mutate(term = reorder_within(term, distinctivity, topic)) %>%
  ggplot(aes(distinctivity, term, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") +
  scale_y_reordered()
```

Summary: 

Topic 1 is likely focused on financial markets and economic indicators, discussing stock prices, market trends, and financial terms such as billion, million, dollar, percent changes, and new developments.

Topic 2 seems to center around legislative activities, likely in the United States, involving the passing of bills, budget considerations, trade discussions, and perhaps the involvement of President Bush, indicating a specific historical timeframe.

Topic 3 appears to discuss geopolitical matters, particularly relating to the Middle East, with references to Iraq, Kuwait, Saudi Arabia, and issues like war, oil, and state affairs.

Topic 4 seems to be related to security and military operations, with words indicating law enforcement, military engagement, and possibly conflict or violence.

Topic 5 might pertain to political dynamics, discussing government parties, presidential matters, national issues, and perhaps focusing on South Africa or southern regions in a broader context.

Topic 6 appears to deal with domestic issues or public safety, including police matters, incidents involving individuals (possibly of varying ages), state-level concerns, and hospital or health-related topics.

Topic 7 is centered on business and corporate matters, including company news, financial figures such as million and billion, corporate strategies, and leadership.

Topic 8 focuses on political campaigns, likely in the U.S., referencing figures such as Bush, Dukakis, and Jackson, and involving the Republican and Democratic parties, suggesting an election context.

Topic 9 seems to be about air travel or aviation, with references to news about flights, alignment (possibly referring to airline alliances or routes), and mentions of a president, which could be related to aviation policies or incidents.

Topic 10 covers topics related to the Cold War era, with terms related to the Soviet Union, East and West Germany, and possibly the policies or actions of President Gorbachev.

Topic 11 deals with legal proceedings and law enforcement, discussing federal courts, legal cases, the roles of attorneys and judges, state laws, and possibly drug-related charges.

Topic 12 appears to relate to temporal events or societal trends, with a focus on new developments, work, time management, and perhaps reflecting on the past year or looking forward to the next.
